RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-09-14 gmas.tex
Finished section 4 (for now)

The objective of our case study was to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks. In order to study this, we manually coded transcripts of such discussions. For our coding, we used a list of argument schemes and critical questions based on GRL and PRAS. In this section we present our case study. All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\url{insertURL}}.

@@

\begin{table}[t]
\centering
\begin{tabularx}{0.5\textwidth}{|l|X|l|l|l|>{\bfseries}l|}
\hline
\multicolumn{2}{|c|}{\textbf{Scheme/Question}} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
\hline 
AS0 & Actor & 2 & 2 & 5 & 9\\
\hline
AS1 & Resource & 2 & 4 & 5 & 11\\
\hline
AS2 & Task/action & 20 & 21 & 17 & 58\\
\hline
AS3 & Goal & 0 & 2 & 2 & 4\\
\hline
AS4 & Softgoal & 3 & 4 & 2 & 9\\
\hline
AS5 & Goal decomposes into tasks & 4 &0& 4 & 8\\
\hline
AS6 & Task contributes (negatively) to softgoal & 8 & 3 &0& 11\\
\hline
AS7 & Goal contributes (negatively) to softgoal &0& 1 & 1 & 2\\
\hline
AS8 & Resource contributes to task & 0 & 4 & 3 & 7\\
\hline
AS9 & Actor depends on actor &0& 1 & 3 & 4\\
\hline
AS10 & Task decomposes into tasks & 11 &14 &11 &36\\ 
\hline
\hline
CQ2 & Task is possible? & 2 & 2 & 1 & 5\\
\hline		
CQ5a & Does the goal decompose into the tasks? & 0 & 1 & 0 & 1\\
\hline
CQ5b & Goal decomposes into other tasks? & 1 & 0 & 0 & 1\\
\hline
CQ6b & Task has negative side effects? & 2 & 0 & 0 & 2\\
\hline
CQ10a & Task decompose into other tasks? & 1 &2 &0&3\\
\hline
CQ10b & Decomposition type correct? &1 &0& 1 &2\\
\hline
\hline
CQ12 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
\hline
CQ13 & Is the element clear/unambiguous? &3 &10 & 3 & 16\\
\hline
\hline
- & Generic counterargument & 0& 2 & 2 & 4\\
\hline
\hline
\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&222\\
\hline
\end{tabularx}
\caption{Occurrences of argument schemes and critical questions in the transcripts.\todo{F}{F}{check numbers against original annotation docs}}
\label{table:transcripts:results:argumentschemes}
\end{table}

@@

>>>>

The objective of our case study was to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks. In order to study this, we manually coded transcripts of such discussions. For our coding, we used a list of argument schemes and critical questions based on GRL and PRAS. In this section we present our case study. All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\todo{F}{M}{\url{insertURL}}}.

@@

\begin{table}[t]
\centering
\begin{tabularx}{0.5\textwidth}{|l|X|l|l|l|>{\bfseries}l|}
\hline
\multicolumn{2}{|c|}{\textbf{Scheme/Question}} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
\hline 
AS0 & Actor & 2 & 2 & 5 & 9\\
\hline
AS1 & Resource & 2 & 4 & 5 & 11\\
\hline
AS2 & Task/action & 20 & 21 & 17 & 58\\
\hline
AS3 & Goal & 0 & 2 & 2 & 4\\
\hline
AS4 & Softgoal & 3 & 4 & 2 & 9\\
\hline
AS5 & Goal decomposes into tasks & 4 &0& 4 & 8\\
\hline
AS6 & Task contributes (negatively) to softgoal & 8 & 3 &0& 11\\
\hline
AS7 & Goal contributes (negatively) to softgoal &0& 1 & 1 & 2\\
\hline
AS8 & Resource contributes to task & 0 & 4 & 3 & 7\\
\hline
AS9 & Actor depends on actor &0& 1 & 3 & 4\\
\hline
AS10 & Task decomposes into tasks & 11 &14 &11 &36\\ 
\hline
\hline
CQ2 & Task is possible? & 2 & 2 & 1 & 5\\
\hline		
CQ5a & Does the goal decompose into the tasks? & 0 & 1 & 0 & 1\\
\hline
CQ5b & Goal decomposes into other tasks? & 1 & 0 & 0 & 1\\
\hline
CQ6b & Task has negative side effects? & 2 & 0 & 0 & 2\\
\hline
CQ10a & Task decompose into other tasks? & 1 &2 &0&3\\
\hline
CQ10b & Decomposition type correct? &1 &0& 1 &2\\
\hline
\hline
CQ11 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
\hline
CQ12 & Is the element clear/unambiguous? &3 &10 & 3 & 16\\
\hline
\hline
- & Generic counterargument & 0& 2 & 2 & 4\\
\hline
\hline
\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&222\\
\hline
\end{tabularx}
\caption{Occurrences of argument schemes and critical questions in the transcripts.\todo{F}{F}{check numbers against original annotation docs}}
\label{table:transcripts:results:argumentschemes}
\end{table}

@@

Our coding further led us to identify three different operations, that is, different effects an argument or critical question can have on a goal model: an argument can introduce a new goal model element (\textsf{INTRO}); it can disable (i.e. attack) a goal model element (\textsf{DISABLE}); or it can replace a goal model element (\textsf{REPLACE}). Consider, for example, Table~\ref{table:transcripts:traffic-light}. First, an argument is posed that introduces a number of tasks. Counterarguments are then given against two of these tasks (\emph{add traffic light} and \emph{remove intersection}), which are subsequently disabled. An example of \textsf{REPLACE} is given in Table~\ref{table:transcript:decomposition}: what used to be an AND-decomposition is changed into an OR-decomposition. We will further discuss the possible operations in Section~\ref{sect:overview}.
